Impact ionization processes in a photodriven Mott insulator: influence of phononic dissipation

### 2022-08-26 Paper1.tex
Final version 26.08.2022

We study a model for photovoltaic energy collection consisting of a Mott insulating layer in presence of acoustic phonons, coupled to two wide-band fermion leads at different chemical potentials and driven %PGCH[
into a nonequilibrium steady state
%PGCH]
by a periodic electric field. We treat electron correlations with nonequilibrium dynamical mean-field theory (DMFT) using the so-called auxiliary master equation approach as impurity solver and include dissipation by acoustic phonons via the Migdal approximation. For a small hybridization to the leads, we obtain a peak in the photocurrent as a function of the driving frequency which can be associated with impact ionization processes.
For larger hybridizations the shallow peak suggests a suppression of impact ionization with respect to direct photovoltaic excitations. Acoustic phonons slightly enhance the photocurrent for small driving frequencies and suppress it at frequencies around the main peak at all considered hybridization strengths.

@@

%PGCH[ I changed some words and rephrased the last sentence in this paragraph after Enrico modified it

@@

%PGCH]

@@

%PGCH: I put this section as a subsection as before

@@

For $\Omega\gtrsim8$, the current increases further until it reaches its peak at $\Omega\approx 11$.
In this range both DE and II can occur in principle. The height of the peak is approximately the same for both $v$, despite
%PGCH: REMOVED:the lead-layer hybridizations differing by a factor of two. ALTERNATIVE FORMULATION:
%PGCH[
they differ by a factor of two.
%PGCH]
The double occupation in this range shows a similar behavior as the current for $v=0.4$ with a peak around $\Omega\approx 11$, while it exhibits a plateau for $v=0.8$.
For $v=0.4$ the current curve is characterized by a substantial increase in slope around $\Omega \approx 7$, which cannot be observed for $v=0.8$. This, together with the comparable magnitude of the
%PGCH[
current
%PGCH]
maxima at $\Omega\approx 11$ and the behavior of the double occupation strongly suggests that for $v=0.4$ a substantial rate of II processes take place, while these are absent or negligible for $v=0.8$.
Here,  the plateau in $N_{\text{D}}$ for $v=0.8$ suggests that photoexcited electrons arriving in the UHB quickly escape to the right fermionic lead and do not have time to induce II processes.

@@

%PGCH[ Added Enrico's part. I changed some words.

@@

%PGCH]

%PGCH: I put this section as a subsection as before

@@

The double occupation $N_{\text{D}}$  shown in Fig.~\ref{fig:j_vs_omega_mu1_v_0.4_0.8_eph}(b) is larger than and follows the behavior of its EO counterpart within the entire $\Omega$-range considered.
The spectral and occupation functions  shown in Fig.~\ref{fig:spec_filling_mu1_v_0.4_0.8_O_11_eph} display a redistribution of spectral weight from the peaks towards the edges of the bands, thus reducing the pseudogap. Consequently, the occupation of positive energy states is slightly shifted to the bottom of the UHB compared to the EO case, as the occupation function shows.
 The spectral function in presence of phonons is  broadened and tends to fill the pseudogap $\Delta_{\text{pg}}$ as anticipated in Sec.~\ref{sec:scheme_electrons_phonons}.
This reduction of the pseudogap explains why the current in Fig.~\ref{fig:j_vs_omega_mu1_v_0.4_0.8_eph}(a) is slightly larger in presence of phonons for $\Omega\lesssim7$. %PGCH: removed reference to in gap phonon states
With a smaller pseudogap and more states at its edges, more electrons can be directly excited with a given driving frequency $\Omega$ thereby increasing the photocurrent. On the other hand, the suppression of the current by e-ph interaction in the range $7\lesssim\Omega\lesssim14$ is due to the dissipation-induced reduction of spectral weight at frequencies $|\omega| \gtrsim 3.5$, cf. Fig. \ref{fig:spec_filling_mu1_v_0.4_0.8_O_11_eph}.

@@

Summarizing, phonons
%PGCH[ 
slightly
%PGCH]
enhance the current for almost all driving frequencies $\Omega\lesssim 7$ for which only DEs take place. They suppress the current in the high-frequency range $7\lesssim\Omega\lesssim14$.

@@

The results above suggest that the influence of acoustic phonons on the electronic scattering processes does not depend significantly on $v$, i.e. on whether II processes are significant or not. In other words, the effect of phonons doesn't depend significantly on the rate at which carriers are injected into and removed from the layer. Fig.~\ref{fig:j_vs_omega_mu1_sweep_v_E0_2_eph} confirms this behavior for intermediate lead-layer hybridizations $v=0.5,0.6,0.7$. Also the location of the crossing in the current  between the EO and e-ph case is essentially independent of $v$, as can be seen in 
 Fig.~\ref{fig:j_vs_omega_mu1_sweep_v_E0_2_eph}(a).
%PGCH: I find this sentence below not clearly motivated. In the e-ph case, the correlation between the increase of N_D and the current suppression is true, but it's also true that the double occupation is larger than the EO counterpart also when the current with phonons is enhanced in the low-frequency range.   
The increase of double occupation in the case with e-ph interaction 
(cf. Fig.~\ref{fig:j_vs_omega_mu1_v_0.4_0.8_eph}(b)-Fig.~\ref{fig:j_vs_omega_mu1_sweep_v_E0_2_eph}(b))
coincides with the current suppression and is probably due to a suppression of the mobility of photoexcited electrons in the UHB.

@@

%PGCH: Removed the paragraph related to suggestions and possible experimental realization of our setup.

%PGCH[ Added Enrico's part but I changed some words and rephrased

Experimentally, the electric field amplitude considered in this paper~\footnote{We don't consider smaller electric field amplitudes $E_0$ since DMFT only converges slowly for such a choice of parameters.} is several orders of magnitude larger than that of the sunlight, therefore these results are more relevant for photoexcitations produced by intense laser pulses (see also the discussion in ~\cite{mu.we.18}) rather than for true photovoltaic systems. On the other hand, one may try to qualitatively extrapolate the discussion at the end of Sec.~\ref{sec:E0_2_only_electrons} in the other direction,  namely to smaller electric field amplitudes, for which a smaller hybridization 
%PGCH: REMOVED: is. ALTERNATIVE FORMULATION:
%PGCH[
would be
%PGCH]
necessary in order to sustain II. In a more realistic setup modelling oxide heterostructures~\cite{as.bl.13,pe.be.19}, a regime of effectively small hybridization would be achieved by considering several layers in which photoexcited carriers, separated by an electric field gradient, have more time to induce II processes before escaping into the leads. In order to address the occurrence of II in such setups, also further extensions should be taken into account, such as the inclusion of multiorbital correlated systems as well as impurity scattering. Such topics would be interesting for future studies.

%PGCH]

>>>>

We study a model for photovoltaic energy collection consisting of a Mott insulating layer in presence of acoustic phonons, coupled to two wide-band fermion leads at different chemical potentials and driven
into a nonequilibrium steady state by a periodic electric field. We treat electron correlations with nonequilibrium dynamical mean-field theory (DMFT) using the so-called auxiliary master equation approach as impurity solver and include dissipation by acoustic phonons via the Migdal approximation. For a small hybridization to the leads, we obtain a peak in the photocurrent as a function of the driving frequency which can be associated with impact ionization processes.
For larger hybridizations the shallow peak suggests a suppression of impact ionization with respect to direct photovoltaic excitations. Acoustic phonons slightly enhance the photocurrent for small driving frequencies and suppress it at frequencies around the main peak at all considered hybridization strengths.

@@

@@

@@

@@

For $\Omega\gtrsim8$, the current increases further until it reaches its peak at $\Omega\approx 11$.
In this range both DE and II can occur in principle. The height of the peak is approximately the same for both $v$, despite they differ by a factor of two. The double occupation in this range shows a similar behavior as the current for $v=0.4$ with a peak around $\Omega\approx 11$, while it exhibits a plateau for $v=0.8$. For $v=0.4$ the current curve is characterized by a substantial increase in slope around $\Omega \approx 7$, which cannot be observed for $v=0.8$. This, together with the comparable magnitude of the current maxima at $\Omega\approx 11$ and the behavior of the double occupation strongly suggests that for $v=0.4$ a substantial rate of II processes take place, while these are absent or negligible for $v=0.8$.
Here,  the plateau in $N_{\text{D}}$ for $v=0.8$ suggests that photoexcited electrons arriving in the UHB quickly escape to the right fermionic lead and do not have time to induce II processes.

@@

@@

@@

The double occupation $N_{\text{D}}$  shown in Fig.~\ref{fig:j_vs_omega_mu1_v_0.4_0.8_eph}(b) is larger than and follows the behavior of its EO counterpart within the entire $\Omega$-range considered.
The spectral and occupation functions  shown in Fig.~\ref{fig:spec_filling_mu1_v_0.4_0.8_O_11_eph} display a redistribution of spectral weight from the peaks towards the edges of the bands, thus reducing the pseudogap. Consequently, the occupation of positive energy states is slightly shifted to the bottom of the UHB compared to the EO case, as the occupation function shows.
 The spectral function in presence of phonons is  broadened and tends to fill the pseudogap $\Delta_{\text{pg}}$ as anticipated in Sec.~\ref{sec:scheme_electrons_phonons}.
This reduction of the pseudogap explains why the current in Fig.~\ref{fig:j_vs_omega_mu1_v_0.4_0.8_eph}(a) is slightly larger in presence of phonons for $\Omega\lesssim7$. With a smaller pseudogap and more states at its edges, more electrons can be directly excited with a given driving frequency $\Omega$ thereby increasing the photocurrent. On the other hand, the suppression of the current by e-ph interaction in the range $7\lesssim\Omega\lesssim14$ is due to the dissipation-induced reduction of spectral weight at frequencies $|\omega| \gtrsim 3.5$, cf. Fig. \ref{fig:spec_filling_mu1_v_0.4_0.8_O_11_eph}.

@@

Summarizing, phonons slightly enhance the current for almost all driving frequencies $\Omega\lesssim 7$ for which only DEs take place. They suppress the current in the high-frequency range $7\lesssim\Omega\lesssim14$.

@@

The results above suggest that the influence of acoustic phonons on the electronic scattering processes does not depend significantly on $v$, i.e. on whether II processes are significant or not. In other words, the effect of phonons doesn't depend significantly on the rate at which carriers are injected into and removed from the layer. Fig.~\ref{fig:j_vs_omega_mu1_sweep_v_E0_2_eph} confirms this behavior for intermediate lead-layer hybridizations $v=0.5,0.6,0.7$. Also the location of the crossing in the current  between the EO and e-ph case is essentially independent of $v$, as can be seen in 
 Fig.~\ref{fig:j_vs_omega_mu1_sweep_v_E0_2_eph}(a).

@@

Experimentally, the electric field amplitude considered in this paper~\footnote{We don't consider smaller electric field amplitudes $E_0$ since DMFT only converges slowly for such a choice of parameters.} is several orders of magnitude larger than that of the sunlight, therefore these results are more relevant for photoexcitations produced by intense laser pulses (see also the discussion in ~\cite{mu.we.18}) rather than for true photovoltaic systems. On the other hand, one may try to qualitatively extrapolate the discussion at the end of Sec.~\ref{sec:E0_2_only_electrons} in the other direction,  namely to smaller electric field amplitudes, for which a smaller hybridization would be necessary in order to sustain II. In a more realistic setup modelling oxide heterostructures~\cite{as.bl.13,pe.be.19}, a regime of effectively small hybridization would be achieved by considering several layers in which photoexcited carriers, separated by an electric field gradient, have more time to induce II processes before escaping into the leads. In order to address the occurrence of II in such setup, also further extensions should be taken into account, such as the inclusion of multiorbital correlated systems as well as impurity scattering. Such topics would be interesting for future studies.
